---
title: "Using Matching Subset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{test_subsetdefinitions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(HadesExtras)
devtools::load_all(".")
#options("DEBUG_DATABASECONNECTOR_DBPLYR" = FALSE)
```

# Configuration 

Create a `CohortTableHandles` with a test cohort. 

```{r}
config_yaml <- "
  databaseName: Eunomia1
  connection:
    connectionDetailsSettings:
        dbms: eunomia
  cdm:
      cdmDatabaseSchema: main
      vocabularyDatabaseSchema: main
  cohortTable:
      cohortDatabaseSchema: main
      cohortTableName: test_cohort_table"

cat(config_yaml)
```

```{r}
config <- yaml::yaml.load(config_yaml)
```

```{r}
configurations <- yaml::read_yaml(testthat::test_path("config", "test_config.yml"))
configurationName <- "dev_bigquery"
configurationName <- "dev_eunomia"
config <- configurations[[configurationName]]
```

# Create CohortTableHandler

As in `CDMHandled`, to create a `CohortTableHandles`, for convenience `createCohortTableHandlerFromList` can be used. 
```{r}
cohortTableHandler <- createCohortTableHandlerFromList(config)
```

In addition to the checks performed by `CDMHandled`, `cohortTableHandler` includes a check on the creation of the `cohort` table. 

```{r}
cohortTableHandler$connectionStatusLog |> 
  reactable_connectionStatus()
```

```{r}

# cohort 10:
# 1 M born in 1970
# 1 F born in 1971
#
# cohort 20:
# 10 M born in 1970
# 10 F born in 1970
# 10 F born in 1971
# 10 F born in 1972

cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = here::here("inst/testdata/matching/Cohorts.csv"),
  jsonFolder = here::here("inst/testdata/matching/cohorts"),
  sqlFolder = here::here("inst/testdata/matching/sql/sql_server"),
  cohortFileNameFormat = "%s",
  cohortFileNameValue = c("cohortName"),
  #packageName = "HadesExtras",
  verbose = FALSE
)

```

# Create a matching subset 

```{r}
# Match to sex and bday, match ratio 10
  subsetDef <- CohortGenerator::createCohortSubsetDefinition(
    name = "",
    definitionId = 20,
    subsetOperators = list(
      createMatchingSubset(
        matchToCohortId = 10,
        matchRatio = 10,
        matchSex = TRUE,
        matchBirthYear = TRUE,
        matchCohortStartDateWithInDuration = FALSE,
        newCohortStartDate = "keep",
        newCohortEndDate = "keep"
      )
    )
  )
```

```{r}

cohortDefinitionSetWithSubsetDef <- cohortDefinitionSet |>
  CohortGenerator::addCohortSubsetDefinition(subsetDef, targetCohortIds = 20)

cohortDefinitionSetWithSubsetDef |>  tibble::as_tibble() |> dplyr::select(-sql, -json)
```


Build cohorts 
```{r}
cohortTableHandler$insertOrUpdateCohorts(cohortDefinitionSetWithSubsetDef |> dplyr::filter(subsetDefinitionId==20))
```

```{r}
cohortsSummary <- cohortTableHandler$getCohortsSummary()
```

```{r}
rectable_cohortsSummary(cohortsSummary)
```














