---
title: "Using CohortTableHandler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{connection_tutorial}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(HadesExtras)
devtools::load_all(".")
#options("DEBUG_DATABASECONNECTOR_DBPLYR" = FALSE)
```

# Intro

The CohortTableHandler is an extension of `CDMHandled` (see vignette) to include the a `cohort` table and the functions to work with this table. 

# Configuration 

In addition to the configuration for `CDMHandled`, we also need the configuration indicating the location of the cohort table. 

We need the following configuration. We write this in yaml format for clarity. 

```{r}
config_yaml <- "
  databaseName: Eunomia1
  connection:
    connectionDetailsSettings:
        dbms: eunomia
  cdm:
      cdmDatabaseSchema: main
      vocabularyDatabaseSchema: main
  cohortTable:
      cohortDatabaseSchema: main
      cohortTableName: test_cohort_table"

cat(config_yaml)
```

```{r}
config <- yaml::yaml.load(config_yaml)
```

```{r}
configurations <- yaml::read_yaml(testthat::test_path("config", "test_config.yml"))
configurationName <- "dev_bigquery"
configurationName <- "dev_eunomia"
config <- configurations[[configurationName]]
```

# Create CohortTableHandler

As in `CDMHandled`, to create a `CohortTableHandles`, for convenience `createCohortTableHandlerFromList` can be used. 
```{r}
cohortTableHandler <- createCohortTableHandlerFromList(config)
```

In addition to the checks performed by `CDMHandled`, `cohortTableHandler` includes a check on the creation of the `cohort` table. 

```{r}
cohortTableHandler$connectionStatusLog |> 
  reactable_connectionStatus()
```


# Add cohors to `cohort` table 


To add cohort to the `cohortTableHandler` object we need a `cohortDefinitionSet` table (this table is defined in [CohortGenerator](https://ohdsi.github.io/CohortGenerator/articles/GeneratingCohorts.html#downloading-cohorts-from-atlas)). 

`cohortDefinitionSet` can be created from
[Atlas](https://ohdsi.github.io/CohortGenerator/articles/GeneratingCohorts.html#downloading-cohorts-from-atlas) 
cohort definitions, or from 
[study pacakages](https://ohdsi.github.io/CohortGenerator/reference/getCohortDefinitionSet.html). 

Additionally, `HadesExtras` includes a way to produce `cohortDefinitionSet` based on a list of local person ids (OMOP table person field person_source_value). 

## Add cohorts from Atlas

Get `cohortDefinitionSet` form 3 cohort definitions in Atlas demo. 
```{r}
# webAPI url for the Atlas demo
baseUrl <- "https://api.ohdsi.org/WebAPI"
# A list of cohort IDs for use in this vignette
cohortIds <- c(1778211, 1778212, 1778213)
# Get the SQL/JSON for the cohorts
cohortDefinitionSet <- ROhdsiWebApi::exportCohortDefinitionSet(
  baseUrl = baseUrl,
  cohortIds = cohortIds
)
```

Build cohorts 
```{r}
cohortTableHandler$insertOrUpdateCohorts(cohortDefinitionSet)
```

Check cohort counts 
```{r}
cohortTableHandler$getCohortCounts()
```

```{r}
cohortsSummary <- cohortTableHandler$getCohortsSummary()
```

```{r}
rectable_cohortsSummary(cohortsSummary)
```


## Add cohorts from cohortData file

Create tibble of `cohortData` format. See the documentation for the correct format. 

```{r}
cohortData <- tibble::tribble(
  ~cohort_name, ~person_source_value, ~cohort_start_date, ~cohort_end_date,
  "Cohort A", "000728a7-80de-420a-9286-2c20e81cb7b8", as.Date("2020-01-01"), as.Date("2020-01-03"),
  "Cohort A", "000cb58f-523d-49a2-a05e-de1e93f35c01", as.Date("2020-01-01"), as.Date("2020-01-03"),
  "Cohort A", "001f4a87-70d0-435c-a4b9-1425f6928d33", as.Date("2020-01-01"), as.Date("2020-01-03"),
  "Cohort A", "002805e7-7624-4cb7-b68d-e8ac92f61ff9", as.Date("2020-01-01"), as.Date("2020-01-03"),
  "Cohort A", "0030eb48-316c-4250-907f-a272909ff8b9", as.Date("2020-01-01"), as.Date("2020-01-03"),
  "Cohort B", "00093765-abef-4a56-9280-8f92422afae7", as.Date("2020-01-01"), as.Date("2020-01-04"),
  "Cohort B", "00196a95-1567-41f8-b608-18e6295b4c1e", as.Date("2020-01-01"), as.Date("2020-01-04"),
  "Cohort B", "00211cd2-f171-45d9-a7c6-ea8ac6d28d09", as.Date("2020-01-01"), as.Date("2020-01-04"),
  "Cohort B", "0029df47-1263-4576-8ca3-4615adb7dd7a", as.Date("2020-01-01"), as.Date("2020-01-04"),
  "Cohort B", "00444703-f2c9-45c9-a247-f6317a43a929", as.Date("2020-01-01"), as.Date("2020-01-04")
)
```

We can check if the format is correct using `checkCohortData`. See the documentation for the checks

If correct, It will return TRUE. 
```{r}
checkCohortData(cohortData)
```

If any error, It will return an array of strings with the errors detected.  
```{r}
cohortDataWithErrors <- cohortData
cohortDataWithErrors[1,1] <- as.character(NA)
cohortDataWithErrors[1,3] <- as.Date("2030-01-01")
checkCohortData(cohortDataWithErrors)
```


To copy the `cohortData` into the `cohort`it has to be converted to a `cohortDefinitionSet`.  


```{r}
cohortDefinitionSet <- cohortDataToCohortDefinitionSet(
    cohortData = cohortData
  )
```


Copy `cohortDefinitionSet` to `cohort` table 

```{r}
cohortTableHandler$insertOrUpdateCohorts(cohortDefinitionSet)
```


```{r}
cohortsSummary <- cohortTableHandler$getCohortsSummary()
```

```{r}
rectable_cohortsSummary(cohortsSummary)
```





