---
title: "Using TimeCodeWAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{test_subsetdefinitions}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(HadesExtras)
devtools::load_all(".")
#options("DEBUG_DATABASECONNECTOR_DBPLYR" = FALSE)
```

# Configuration 

Create a `CohortTableHandles` with a test cohort. 

```{r}
config_yaml <- "
  databaseName: Eunomia1
  connection:
    connectionDetailsSettings:
        dbms: eunomia
  cdm:
      cdmDatabaseSchema: main
      vocabularyDatabaseSchema: main
  cohortTable:
      cohortDatabaseSchema: main
      cohortTableName: test_cohort_table"

cat(config_yaml)
```

```{r}
config <- yaml::yaml.load(config_yaml)
```

```{r}
configurations <- yaml::read_yaml(testthat::test_path("config", "test_config.yml"))
configurationName <- "dev_bigquery"
configurationName <- "dev_eunomia"
config <- configurations[[configurationName]]
```

# Create CohortTableHandler

As in `CDMHandled`, to create a `CohortTableHandles`, for convenience `createCohortTableHandlerFromList` can be used. 
```{r}
cohortTableHandler <- createCohortTableHandlerFromList(config)
```

In addition to the checks performed by `CDMHandled`, `cohortTableHandler` includes a check on the creation of the `cohort` table. 

```{r}
cohortTableHandler$connectionStatusLog |> 
  reactable_connectionStatus()
```

# Import cohorts

```{r eval=FALSE}
cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = "testdata/asthma/Cohorts.csv",
  jsonFolder = "testdata/asthma/cohorts",
  sqlFolder = "testdata/asthma/sql/sql_server",
  cohortFileNameFormat = "%s",
  cohortFileNameValue = c("cohortId"),
  subsetJsonFolder = "testdata/asthma/cohort_subset_definitions/",
  packageName = "HadesExtras",
  verbose = FALSE
) |> tibble::as_tibble()

```


```{r}
cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = here::here("inst/testdata/asthma/Cohorts.csv"),
  jsonFolder = here::here("inst/testdata/asthma/cohorts"),
  sqlFolder = here::here("inst/testdata/asthma/sql/sql_server"),
  cohortFileNameFormat = "%s",
  cohortFileNameValue = c("cohortId"),
  subsetJsonFolder = here::here("inst/testdata/asthma/cohort_subset_definitions/"),
  #packageName = "HadesExtras",
  verbose = FALSE
) |> tibble::as_tibble()

```

# Create a matching subset 

```{r}
# Match to sex and bday, match ratio 10
  subsetDef <- CohortGenerator::createCohortSubsetDefinition(
    name = "",
    definitionId = 1,
    subsetOperators = list(
      createMatchingSubset(
        matchToCohortId = 1,
        matchRatio = 10,
        matchSex = TRUE,
        matchBirthYear = TRUE,
        matchCohortStartDateWithInDuration = FALSE,
        newCohortStartDate = "asMatch",
        newCohortEndDate = "keep"
      )
    )
  )
```

```{r}

cohortDefinitionSetWithSubsetDef <- cohortDefinitionSet |>
  CohortGenerator::addCohortSubsetDefinition(subsetDef, targetCohortIds = 2)

cohortDefinitionSetWithSubsetDef |>  tibble::as_tibble() |> dplyr::select(-sql, -json)
```



Build cohorts 
```{r}
cohortTableHandler$insertOrUpdateCohorts(cohortDefinitionSetWithSubsetDef)
```

```{r}
cohortTableHandler$getCohortsSummary() |> 
  rectable_cohortsSummary()
```


# calcualte table 

```{r}
temporalStartDays = c(-365*5,   -365*4,   -365*3,   -365*2, -365*1, -3,   4,   365+1, 365*2+1, 365*3+1, 365*4+1)
temporalEndDays =   c(-365*4-1, -365*3-1, -365*2-1, -365*1-1,   -4,  3, 365,   365*2, 365*3,   365*4,   365*5)

temporalCovariateSettings = FeatureExtraction::createTemporalCovariateSettings(
  useConditionOccurrence = TRUE, 
  useDrugExposure =  TRUE, 
  useProcedureOccurrence = TRUE, 
  useMeasurement = TRUE,
  useObservation = TRUE,
  temporalStartDays = temporalStartDays, 
  temporalEndDays =   temporalEndDays
)

```

```{r}

timeCodeWasCounts <- CohortDiagnostics_runTimeCodeWAS(
    connection = cohortTableHandler$connectionHandler$getConnection(),
    cdmDatabaseSchema = cohortTableHandler$cdmDatabaseSchema,
    vocabularyDatabaseSchema = cohortTableHandler$vocabularyDatabaseSchema,
    cohortDatabaseSchema = cohortTableHandler$cohortDatabaseSchema,
    cohortTable = cohortTableHandler$cohortTableNames$cohortTable,
    cohortIdCases = 1,
    cohortIdControls = 2001,
    covariateSettings = temporalCovariateSettings
)

```

```{r}
timeCodeWasCounts
```




