---
title: "Using CDMHandler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{connection_tutorial}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(HadesExtras)
devtools::load_all(".")
#options("DEBUG_DATABASECONNECTOR_DBPLYR" = FALSE)
```

# Configuration 

We need the following configuration. We write this in yaml format for clarity. 

```{r}
config_yaml <- "
  connection:
    connectionDetailsSettings:
        dbms: eunomia
  cdm:
      cdmDatabaseSchema: main
      vocabularyDatabaseSchema: main
"

cat(config_yaml)
```

```{r}
config <- yaml::yaml.load(config_yaml)
```

```{r}
configurations <- yaml::read_yaml(testthat::test_path("config", "test_config.yml"))
configurationName <- "dev_bigquery"
configurationName <- "dev_eunomia"
config <- configurations[[configurationName]]
```

# CDMHandled

CDMHandled is a R6 object connected to a OMOP-CDM. 
It includes attributes and functions to retrieve information from the database and work with the tables.

```{r, eval=FALSE}
connectionHandler <- ResultModelManager_createConnectionHandler(
    connectionDetailsSettings = config$connection$connectionDetailsSettings,
    tempEmulationSchema = config$connection$tempEmulationSchema, 
    useBigrqueryUpload = config$connection$useBigrqueryUpload
  )

  CDMdb <- CDMdbHandler$new(
    databaseName = config$databaseName,
    connectionHandler = connectionHandler,
    cdmDatabaseSchema = config$cdm$cdmDatabaseSchema,
    vocabularyDatabaseSchema = config$cdm$vocabularyDatabaseSchema
  )
```

Or for convenience `createCDMdbHandlerFromList` can be used. 
```{r}
CDMdb <- createCDMdbHandlerFromList(config)
```


First thing to do is to check the status of the connection. 
`CDMdb$connectionStatusLog` not only shows if the connection is correct, but also if the `tempEmulationSchema` works, and all the CDM and Vocabulary tables are available in the database.  
```{r}
CDMdb$getConnectionStatus()
```

Function `reactable_connectionStatus` creates an interactive table with this information.
```{r}
CDMdb$getConnectionStatus() |> 
  reactable_connectionStatus()
```

We can get information on the CDM and Vocabulary version: 
```{r}
CDMdb$CDMInfo
```

```{r}
CDMdb$vocabularyInfo
```

Tables in the database can be used as if they were tibbles thanks to  [dbplyr](https://dbplyr.tidyverse.org/). 
```{r}
CDMdb$getTblCDMSchema$person() |> 
  dplyr::count(year_of_birth) 
```

```{r}
CDMdb$getTblCDMSchema$person() |>
  dplyr::left_join(
    CDMdb$getTblVocabularySchema$concept() |> dplyr::select(concept_id, concept_name),
    by = c("gender_concept_id" = "concept_id")
  ) |> 
  ggplot2::ggplot(ggplot2::aes(x=year_of_birth, fill=factor(concept_name))) + 
    ggplot2::geom_bar()
```








